---
title: Preproviz
author: Markus Vattulainen
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Preproviz}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

Data quality issues such as missing values and outliers are often
interdependent, which makes preprocessing both time-consuming and leads to
suboptimal performance in knowledge discovery tasks. This package supports
preprocessing decision making by visualizing interdependent data quality issues
through means of feature construction. The user can define his own application
domain specific constructed features that express the quality of a data point
such as number of missing values in the point or use nine default features.
The outcome can be explored with plot methods and the feature constructed data
acquired with get methods.

## Quick start

```{r,fig.width=8, fig.height=8, message = FALSE, warnings = FALSE}
library(preproviz)
result <- preproviz(iris)
plotDENSITY(result)
```

## Comparisons

The package supports comparison of multiple data sets or different versions of a same data set.

Let's make some test data.

```{r, message = FALSE, warnings = FALSE}
iris2 <- iris
iris2[sample(1:150,30), 1] <- NA # adding missing values
iris2[sample(1:150,30), 5] <- levels(iris2$Species)[2] # adding inconsistency 
```

and then setup comparison between iris  and iris2.

```{r,fig.width=8, fig.height=8, message = FALSE, warnings = FALSE}
result <- preproviz(list(iris, iris2))
```

The results can be plotted with several plot methods. First, plotting how the constructed features cluster (that is, which features are
linearly dependent on each other). 

```{r,fig.width=8, fig.height=8, message = FALSE, warnings = FALSE}
plotVARCLUST(result)
```

and then how the constructed feature data points cluster when reduced
to two dimensions.

```{r,fig.width=8, fig.height=8, message = FALSE, warnings = FALSE}
plotCMDS(result)
```

## Customization

The setups can be customized:

```{r, message = FALSE, warnings = FALSE, eval=FALSE}
customparameters <- initializeparameterclassobject(list("LOFScore", "ScatterCounter"))
setup1 <- initializesetupclassobject("setup1", customparameters, initializedataobject(iris))
setup2 <- initializesetupclassobject("setup2", customparameters, initializedataobject(otherdataframehere)) 
control <- initializecontrolclassobject(list("setup1", "setup2")) 
result <- preproviz(control)
```

and new constructe features added to the system:

```{r, message = FALSE, warnings = FALSE, eval=FALSE}
constructfeature("MissingValueShare", "apply(data, 1, function(x) sum(is.na(x))/ncol(data))", impute=TRUE)
```

## Default contructed features

There are nine default constructed features:

- MissingValueShare, count the number of missing values on a row and divide it by the total number of features

- LenghtOfIQR, min-max normalize the data and compute the length of IQR for each row (i.e. flatness of reply)

- DistanceToNearest, min-max normalize the data and compute the Euclidean distance to the nearest data point

- ClassificationCentainty, compute the random forest class propability that has the highest value

- ClassificationScatter. compute the scatter counter (number of changes in class labels) in 1/10 point neighbourhood

- NearestPointPurity,Neareast point is of same class or not

- NeighborhoodDiversity, count the number of dominant class points in 1/10 data set size neighborhood and divide 
it with the total number of classes in the data set.

- LOFScore, compute LOF scores

- MahalanobisDistance, compute Mahalanobis distance to class center

- ClusteringTendency, Count the Hopkins statistic without a row and divide it with Hopkings statistic for all rows



